### 2014-11-04 application.tex
More updates.

To better understand how coordination facts are used, we next present some programs that
take advantage of them.

>>>>

To better understand how coordination facts are used, we next present some programs that
take advantage of them. In our experimental setup, we we used a machine with
four AMD Six-Core Opteron TM 8425 HE (2100 MHz) chips (24 cores) and 64 GB of
DDR-2 667MHz (16x4GB) RAM, running GNU/Linux (kernel 3.15.10-201 64 bits).
We compiled our virtual machine using GCC 4.8.3 (g++) with the flags
\texttt{-O3 -std=c++11 -fno-rtti -march=x86-64}~\footnote{Implementation
   available in \url{http://github.com/.../meld}}.

### 2014-11-04 heat-transfer.tex
More updates.

Randomized and approximation
algorithms can obtain significant benefits from coordination directives because although the final
program results will not be exact, they follow important statistical properties and can be computed faster.
Examples of such programs are Loopy Belief
Propagation~\cite{Gonzalez+al:aistats09paraml} (shown in the next section) and
Heat Transfer.

>>>>

Randomized and approximation
algorithms can obtain significant benefits from coordination directives because although the final
program results will not be exact, they follow important statistical properties and can be computed faster.
Examples of such programs are Loopy Belief
Propagation~\cite{Gonzalez+al:aistats09paraml} (shown in the next section) and
Heat Transfer~(HT).

In HT, we have a graph of nodes that exchange heat values between them. The goal
of the program is to exchange heat to a point where new heat value $\delta = |H_i -
H_{i-1}| \le \epsilon$ for all nodes of the graph. The algorithm works
asynchronously since heat values can be updated by using new partial information
coming from neighbor nodes. This increases parallelism since nodes do not need
to synchronize if we had to compute heat values using iterations.

Fig.~\ref{code:ht} shows the HT logical rules that send new heat values to
neighbor nodes. We added the first rule to increase the priority of the neighbor
nodes if the current node has a large $\delta$. The idea is to prioritize the
computation of heat values of nodes (using \texttt{update}) that have a neighbor
node that changed significantly. With \texttt{set-priority} the node is
temporarily prioritized and selected to execute before others.

\begin{figure}[h!]
\scriptsize\begin{Verbatim}[numbers=left,commandchars=*\[\]]
new-heat(A, New, Old),
Delta = fabs(New - Old),
Delta > epsilon * 2.0
   -o {B, W | !edge(A, B, W) |
         new-neighbor-heat(B, A, New),
         update(B), *underline[set-priority(B, Delta)]}.

new-heat(A, New, Old),
Delta = fabs(New - Old)
Delta < epsilon * 2.0,
Delta > epsilon
   -o {B, W | !edge(A, B, W) |
         new-neighbor-heat(B, A, New), update(B)}.

new-heat(A, New, Old)
Delta = fabs(New - Old)
Delta < epsilon
   -o {B, W | !edge(A, B, W) |
         new-neighbor-heat(B, A, New)}.
\end{Verbatim}
  \caption{Coordination code for the Heat Transfer program. We added one rule
     that covers cases where $\Delta$ is at least $2 \epsilon$ in order to
     increase the priority of neighbors nodes. The code logic remains exactly
     the same as before, however bigger changes in heat values are now
     propagated faster.}
  \label{code:ht}
\end{figure}
\normalsize

### 2014-11-04 minimax.tex
More updates.

>>>>

In Fig.~\ref{results:memory_minmax} we compare the memory usage and scalability
of the coordinated MiniMax against the regular MiniMax. The coordinated version
sees a huge reduction of memory usage, needing at most 8MB for 16 threads, while
the regular versio nneeds around 14GB. There is also some
correlation between memory usage and number of threads. This is explained by the
need for more data structures for handling multiple threads, specially
concerning the memory allocator used by the virtual machine. In terms of run
time, our experimental results show linear scalability for the coordinated
version and a 20\% run time reduction when using 16 threads.

### 2014-11-04 splash-bp.tex
More updates.

\begin{description}
   \item[Tree building]: Each node has a \texttt{waiting} fact that is used to
   start the tree building process. When the highest priority is picked, a token
   is created that will navigate through the tree. Note that in the rule located
   in lines 13-20, newly added nodes to the tree must have a positive priority
   (due to new belief updates) and must be residing in the same thread.
   We want the tree to be kept in the same thread in order to maximize
   parallelism with 1 tree per thread.
   \item[First phase]: In the third rule (lines 10-11), when the number of nodes
   of the tree reaches a certain limit, a 
   \texttt{first-phase} is generated in order to update the beliefs of all nodes
   in the tree, starting from the leaves and ending at the root
   As the nodes are updated, an \texttt{update} fact is derived to update
   the belief values (line 34).
   \item[Second phase]: In the second phase, the computation of beliefs is
   performed from the root to the leaves and the belief values are updated a
   second time (line 41).
\end{description}

Note that the \texttt{set-static} action fact is used in line 1 in order to
force all nodes to remain in the original threads, so that each thread has a
well defined area of nodes to build splash trees on.

\begin{figure}[h!]
\scriptsize\begin{Verbatim}[numbers=left,commandchars=*\{\}]
*underline{set-static(A)}. // all nodes are static.

@@

@@

\begin{figure*}[]
   \begin{center}
      \subfloat[]{\includegraphics[width=5cm]{results/system_belief-propagation-400.png}}
      \subfloat[]{\includegraphics[width=5cm]{results/system_splash-bp-400.png}}
      \subfloat[]{\includegraphics[width=5cm]{results/system_improve_belief-propagation-400.png}}
   \end{center}
   \caption{Memory usage and scalability of the regular and coordinated versions
      of MiniMax.}
   \label{results:splash_bp}
\end{figure*}

>>>>

\begin{description}
   \item[Tree building]: Each node has a \texttt{waiting} fact that is used to
   start the tree building process. When the highest priority is picked, a token
   is created that will navigate through the tree. Note that in the rule located
   in lines 14-21, newly added nodes to the tree must have a positive priority
   (due to new belief updates) and must be residing in the same thread.
   We want the tree to be kept in the same thread in order to maximize
   parallelism with 1 tree per thread.
   \item[First phase]: In the third rule (lines 11-12), when the number of nodes
   of the tree reaches a certain limit, a 
   \texttt{first-phase} is generated in order to update the beliefs of all nodes
   in the tree, starting from the leaves and ending at the root
   As the nodes are updated, an \texttt{update} fact is derived to update
   the belief values (line 35).
   \item[Second phase]: In the second phase, the computation of beliefs is
   performed from the root to the leaves and the belief values are updated a
   second time (line 42).
\end{description}

Note that the \texttt{set-static} and \texttt{set-cpu} action fact are used in
line 2 in order to (1) force nodes to stay in the thread and (2) to partition
nodes as a grid of threads. This setups a well defined area of nodes
for threads to build splash trees on.

\begin{figure}[h!]
\scriptsize\begin{Verbatim}[numbers=left,commandchars=*\{\}]
!coord(A, X, Y), start(A)
   -o *underline{set-static(A), set-cpu(A, grid(X, Y))}.

@@

\begin{figure*}[ht!]
   \begin{center}
      \subfloat[]{\includegraphics[width=5cm]{results/system_belief-propagation-400.png}}
      \subfloat[]{\includegraphics[width=5cm]{results/system_splash-bp-400.png}}
      \subfloat[]{\includegraphics[width=5cm]{results/system_improve_belief-propagation-400.png}}
   \end{center}
   \caption{Experimental results for LBP and SBP. Figure (a) shows the
      scalability of LBP for both LM and GraphLab and Figure (b) shows the
      scalability of SBP. Figure (c) presents the
      improvements seen in SBP against LBP, where SBP runs, on average, 1.5 to 2.5
      times faster than LBP.}
   \label{results:splash_bp}
\end{figure*}

@@

We measured the behavior of LBP and SBP for both LM and GraphLab.
Fig.~\ref{results:splash_bp} shows that both systems have very similar behavior
when using a variable number of threads.
Note that GraphLab avoids message passing between nodes since new neighbor
belief updates are changed immediatelly, resulting in slightly better
scalability for GraphLab.
